---
title: "Untitled"
author: '...'
date: "2025-02-27"
output: html_document
---

- Variable selection for final model
- R output for the numerical measures of the final model, Report 1) R2, and 2) p-value of ANOVA.
- Interpret the coefficients
- Discuss challenges, if any, that you encounter while working on the project.


```{r}
pay <- read.csv("cleaned_dataset.csv")
```

select Report.Year, Q1.Female, Q2.Female, Q3.Female, Q4.Female, to explain Mean.Hourly.Gap
```{r}
fullmodel <- lm(Mean.Hourly.Gap ~ Report.Year+Q1.Female+Q2.Female+Q3.Female+Q4.Female, data = pay)
summary(fullmodel)
```
From the full model, we derive the equation: \
Mean.Hourly.Gap = -0.087(Report.Year) + 0.35(Q1.Female) + 0.255(Q2.Female) + 0.153(Q3.Female) + -0.731(Q4.Female) \
46.15% of the variation in mean hourly gap can be explained by the variables: Report.Year, Q1.Female, Q2.Female, Q3.Female, Q4.Female
The p-value is 2.2e-16 which is less than 0.05, giving significant evidence to reject the null hypothesis meaning of the predictors in the model is significant.
```{r}
par(mfrow = c(2,2))
plot(fullmodel)
```
Residuals has mean 0
The Q-Q plot shows deviation from normality.
Standardized residuals show non-constant variance
No outliers in Residuals vs Leverage